MixEth: efficient, trustless coin mixing service for Ethereum

### 2018-12-05 article/main.tex
Added Aztec protocol to related work and updated our contribution

The two major techniques to provide mixing services for Ethereum are Möbius, a ring-signature-based solution \cite{meiklejohn2018mobius} and Miximus, a zkSNARK-based proposal \cite{miximus2018}. Both of them burn tremendous amounts of gas to withdraw funds, which could be prohibitive for many use cases. Möbius requires $\num[group-separator={,}]{335714}n$ gas ($n$ is the ring size) while Miximus consumes $\num[group-separator={,}]{1903305}$ gas to verify a zkSNARK proof \cite{miximus2018gascost}. As the Ethereum network is congested, ie. blocks were full during 2018\footnote{https://etherscan.io/chart/gasused}, we argue that it is essential for the network scalability to aim to create protocols and applications that burn as few gas as possible.

\textbf{Our contributions.} In this paper, we present MixEth, a privacy-enhancing protocol and a practical tool for Ethereum, to overcome the above mentioned efficiency issues while retaining strong notions of anonymity, mixer availability and theft prevention already achieved by previous proposals. MixEth requires as few off-chain messages and on-chain transactions as Möbius and Miximus, meanwhile it burns significantly less gas.

@@

In our first implementation of the MixEth protocol, we assumed that all transactions will take place on-chain. Since on-chain storage is extremely expensive, we decided to store only the last two sets of shuffled public keys. Note that storing only the latest shuffle would not be enough, since honest receivers could not prove to the contract that their shuffled public key is compromised unless also the last but first shuffled key is also available for the contract. Such a storage structure implies that after uploading the new shuffled keys, a challenging period should proceed in order to let receivers check the correctness of the shuffle. Furthermore we also allow senders to shuffle and deposit new public keys at the same time, meaning that only $3$ on-chain transactions are sufficient to complete the protocol. However shufflers need to withdraw their shuffling deposits with an additional on-chain transaction. We left it as a future work to determine the optimal ratio of mixed coins and shuffling deposits in order to incentivise correct shuffling.

A great advantage of the fully on-chain version of MixEth is that it allows dynamic anonymity sets. One could potentially deposit funds to the contract and shuffle public keys and leave funds in the mixing contract for indefinite amount of time. As soon as the anonymity set is large enough a receiver could withdraw their assets. The whole process is represented as a finite automata as shown in Figure \ref{fig:shuffleandchallenge}. A receiver in a MixEth contract with $N$ senders could withdraw their funds after $N^{'}$ shuffling rounds, where $N^{'}$ is arbitrary.  This dynamic nature of the contract could even lead to a single monolithic MixEth contract instead of having multiple MixEth contracts with significantly fragmented anonymity sets. However note that the gas complexity of shuffling transactions grow linearly in the number of participants, therefore the fully on-chain implementation is not capable to support extremely large anonymity sets with participants more than cca. $300$.

@@

\begin{table}[H] 
	\caption{Proof-of-concept implementation gas cost results. Expect further improvements. MixEthChannel refers to the implementation which leverages state channels for shuffling and challenging periods}
	\centering 
	\begin{tabular}{@{\extracolsep{6pt}}lccccc@{}} 	
		\toprule
		\hline
		&Deployment&Deposit&\multicolumn{2}{c}{Shuffle}& Withdraw\\
		\cline{4-5}&&&Shuffle upload&Challenge\\
		\hline
		\midrule
		Möbius \cite{meiklejohn2018mobius}     &$\num[group-separator={,}]{1046027}$&$\num[group-separator={,}]{76123}$&0&0&$\num[group-separator={,}]{335714}$n   \\
		Miximus \cite{miximus2018}&$\num[group-separator={,}]{1751378}$&$\num[group-separator={,}]{732815}$&0&0 &$\num[group-separator={,}]{1903305}$  \\
		MixEth&TBD&\num[group-separator={,}]{99254}&\num[group-separator={,}]{138653}+\num[group-separator={,}]{40000}n&\num[group-separator={,}]{227563}&\num[group-separator={,}]{113265}  \\
		MixEthChannel&TBD&TBD&0&0&TBD\\
		\bottomrule
	\end{tabular}
	\label{table:securityproperties}
\end{table}

@@

>>>>

There is no doubt that there exists a tremendous need for privacy overlays for Ethereum as new tools for transaction deanonymization are getting developed and used \cite{chan2017ethereum}. This need of the Ethereum community for privacy was spectacularly embodied in September, 2017 when for several days $68\%$ of all the transaction volume was controlled by a centralized coin mixing service \cite{hugemixer2017}.

The two major techniques to provide decentralized mixing services for Ethereum are Möbius, a ring-signature-based solution \cite{meiklejohn2018mobius} and Miximus, a zkSNARK-based proposal \cite{miximus2018}. Both of them burn tremendous amounts of gas to withdraw funds, which could be prohibitive for many use cases. Möbius requires $\num[group-separator={,}]{335714}n$ gas ($n$ is the ring size) while Miximus consumes $\num[group-separator={,}]{1903305}$ gas to verify a zkSNARK proof \cite{miximus2018gascost}. As the Ethereum network is congested, ie. blocks were full during 2018\footnote{https://etherscan.io/chart/gasused}, we argue that it is essential for the network scalability to aim to create protocols and applications that burn as few gas as possible.

Eventhough (Ethereum) users and transactions can be deanonymized already on the network layer \cite{ethmacrofag2016}, we consider network anonymity an orthogonal problem to that of anonymity on the transaction ledger.

\textbf{Our contributions.} In this paper, we present an efficient mixing protocol for Turing-complete blockchains. This protocol can be used to shuffle cyphertexts, ballots or public keys. The protocol have many usecases: shuffling ElGamal-cyphertexts, decentralized mixnets, e-voting.

To show the practicality of the protocol, we introduce MixEth, a privacy-enhancing protocol and a practical tool for Ethereum, to overcome the above mentioned efficiency issues of Ethereum-based coin-mixers while retaining strong notions of anonymity, mixer availability and theft prevention already achieved by previous proposals (\cite{meiklejohn2018mobius},\cite{miximus2018}). MixEth requires as few off-chain messages and on-chain transactions as Möbius and Miximus, meanwhile it burns significantly less gas.

@@

In our first implementation of the MixEth protocol, we assumed that all transactions will take place on-chain. Since on-chain storage is extremely expensive ($\num[group-separator={,}]{20000}$ gas to store a $256$-bit number), we decided to store only the last two cryptographic accumulators of shuffled public keys. Note that storing only the latest shuffle's accumulator would not be enough, since honest receivers could not prove to the contract that their shuffled public key is compromised unless also the last but first shuffled key is also available for the contract to check its set-membership. Such a storage structure implies that after logging the new shuffled keys and uploading the updated accumulator, a challenging period should proceed in order to let receivers check the correctness of the shuffle and whether their shuffled public key is contained in the accumulator. Furthermore we also allow senders to shuffle and deposit new public keys at the same time, meaning that only $3$ on-chain transactions are sufficient to complete the protocol. However shufflers need to withdraw their shuffling deposits with an additional on-chain transaction. We left it as a future work to determine the optimal ratio of mixed coins and shuffling deposits in order to incentivise correct shuffling.

We chose to implement a pairing-based cryptographic accumulator \cite{catalano2013vector}, since it allows constant-size updates and deletes to the accumulator, while also the set-membership proof is constant-size.

A great advantage of the fully on-chain version of MixEth is that it allows dynamic anonymity sets. One could potentially deposit funds to the contract and shuffle public keys and leave funds in the mixing contract for indefinite amount of time. As soon as the anonymity set is large enough a receiver could withdraw their assets. The whole process is represented as a finite automata as shown in Figure \ref{fig:shuffleandchallenge}. A receiver in a MixEth contract with $N$ senders could withdraw their funds after $N^{'}$ shuffling rounds, where $N^{'}$ is arbitrary.  This dynamic nature of the contract could even lead to a single monolithic MixEth contract instead of having multiple MixEth contracts with significantly fragmented anonymity sets. However note that the gas complexity of shuffling transactions grow linearly in the number of participants, therefore the fully on-chain implementation is not capable to support extremely large anonymity sets with participants more than cca. $\num[group-separator={,}]{15000}$.

@@

\begin{table}[H] 
	\caption{Proof-of-concept implementation gas cost results. Expect further improvements. MixEthChannel refers to the implementation which leverages state channels for shuffling and challenging periods}
	\centering 
	\begin{tabular}{@{\extracolsep{6pt}}lccccc@{}} 	
		\toprule
		\hline
		&Deployment&Deposit&\multicolumn{2}{c}{Shuffle}& Withdraw\\
		\cline{4-5}&&&Shuffle upload&Challenge\\
		\hline
		\midrule
		Möbius \cite{meiklejohn2018mobius}     &$\num[group-separator={,}]{1046027}$&$\num[group-separator={,}]{76123}$&0&0&$\num[group-separator={,}]{335714}$n   \\
		Miximus \cite{miximus2018}&$\num[group-separator={,}]{1751378}$&$\num[group-separator={,}]{732815}$&0&0 &$\num[group-separator={,}]{1903305}$  \\
		MixEth&TBD&$\num[group-separator={,}]{99254}$&$\num[group-separator={,}]{138653}+\num[group-separator={,}]{512}n$&$\num[group-separator={,}]{227563}$&$\num[group-separator={,}]{113265}$  \\
		MixEthChannel&TBD&TBD&0&0&TBD\\
		\bottomrule
	\end{tabular}
	\label{table:securityproperties}
\end{table}

@@

\subsection{Additional related work}
Recently, in November 26, 2018 a new privacy-enhancing protocol and tool was released, called Aztec protocol \cite{aztecprotocol2018}. Aztec allows its users to convert any type of existing digital asset on top Ethereum into Aztec notes. Later on Aztec notes can be traded in a join-split style confidential transaction, namely the value of the traded Aztec notes are encrypted and not visible to outsiders. A join-split style transaction can have multiple inputs and outputs. Aztec applies zero-knowledge proofs and range proofs in order to achieve that the sum of inputs traded in a confidential transaction is greater or equal than the sum of outputs. The gas cost of a confidential Aztec transaction is $\num[group-separator={,}]{219000}$+$\num[group-separator={,}]{121000}n$+$\num[group-separator={,}]{41000}m$, where $n$ and $m$ are the numbers of inputs and outputs respectively.  Although Aztec is not a coin mixer, it can be used not only to provide confidential transaction, but also to increase anonymity. On the downside Aztec uses a trusted setup and considerably more expensive than MixEth to provide the same level of k-anonymity, however it is able to conceal transaction volumes. Most likely a combination of MixEth and Aztec would yield the most gas-efficient solution to achieve k-anonym and confidential transactions. For instance on top Aztec one could use MixEth to shuffle Aztec notes.
